---
title: "Eyewatch Ballarat facebook posts"
author: "Nicholas Spyrison"
date: "July 2018"
output: 
  html_document:
    self_contained: true
---
```{r options, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 8,
  fig.width = 12,
  fig.align = "center",
  cache = FALSE)

require(Rfacebook)
require(tidyverse)
require(plotly)
require(tibble)
require(tsibble)
require(lubridate)

source("./R/pull_fb_posts.r")
source("./R/add_fb_reactions.r") # Joins reactions and saves comments.
source("./R/clean_fb_postreactions.r")
```

Originally a dataset of 952 posts (11 variables) of the Eyewatch Ballarat facebook page were pulled on 16 May 2018.

After cleaning and tidying the data, we removed (105) posts containing 'vicpol' in the link. The follow graphs were created from the remaining 847 posts.

```{r, eval=F,include=F}
# Pull data
### API token has more hoop. quick and dirty manual work:
#load("H:/rfacebook_eyewatch/data/archive/bal_posts_raw.rda")
load("./data/archive/bal_posts_raw.rda")
tmp <- bal_posts_raw

  names(tmp["created_time"]) <- 'created_datetime'
  tmp$created_datetime <- ymd_hms(tmp$created_time)
  tmp$created_date <- as.Date(tmp$created_datetime)
  tmp$created_time <- substr(tmp$created_datetime, 12, 23)
  tmp$created_hour <- as.factor(substr(tmp$created_datetime, 12, 13))
  tmp$created_year <- year(tmp$created_date)
  tmp$created_quarter <- quarter(tmp$created_date, with_year = F)
  tmp$created_month <- month(tmp$created_date, abbr = T)
  tmp$created_week <- week(tmp$created_date)
  tmp$created_yearQuarter <- tsibble::yearquarter(tmp$created_date)
  tmp$created_yearMonth <- tsibble::yearmonth(tmp$created_date)
  tmp$created_yearWeek <- tsibble::yearweek(tmp$created_date)
  tmp$created_weekday <- 
    wday(tmp$created_date, label = TRUE, week_start = 1)
  firstday <- min(tmp$created_date)
  tmp$post_daysfrom0 <- 
    as.integer(tmp$created_date - firstday)
  lastday <- max(tmp$created_date)
  tmp$post_age <- 
    as.integer(lastday - tmp$created_date)
  tmp$type <- as.factor(tmp$type)
      tmp$likes_count <- as.integer(tmp$likes_count)
  # tmp$likes_count.x <- 
  #   as.integer(max(tmp$likes_count.x,tmp$likes_count.y))
  # names(tmp)[names(tmp) == 'likes_count.x'] <- 'likes_count'
  #   tmp <- tmp[names(tmp) != 'likes_count.y']
  tmp$comments_count <- as.integer(tmp$comments_count)
  tmp$shares_count <- as.integer(tmp$shares_count)
  ### NA without reactions.
  # tmp$love_count <- as.integer(tmp$love_count)
  # tmp$haha_count <- as.integer(tmp$haha_count)
  # tmp$wow_count <- as.integer(tmp$wow_count)
  # tmp$sad_count <- as.integer(tmp$sad_count)
  # tmp$angry_count <- as.integer(tmp$angry_count)
  # tmp$reactions_count <- as.integer(tmp$likes_count + 
  #   tmp$love_count + tmp$haha_count + tmp$wow_count + 
  #   tmp$sad_count + tmp$angry_count
  # )
  tmp$link_has_vicpol <- 
    (grepl("vicpol", tmp$link) & tmp$type == "link")
  station <- strsplit(tmp$from_name,' ')[[1]][3]
  tmp$link_is_secure_vicpol <- 
    (grepl("https", tmp$link) & tmp$link_has_vicpol == T)
  
bal_posts_clean <- tmp
save(bal_posts_clean,
     #file="H:/rfacebook_eyewatch/data/archive/bal_posts_clean.rda")
     file="./data/archive/bal_posts_clean.rda")
```

## Posts by time

```{r, echo=F}
#load(file="H:/rfacebook_eyewatch/data/archive/bal_posts_clean.rda")
load(file="./data/archive/bal_posts_clean.rda")

bal_posts_clean <- dplyr::filter(bal_posts_clean, link_has_vicpol == F)
  #952 to 847 obs.

(p1 <- 
    ggplot(data = bal_posts_clean, aes(x = created_date) ) +
    geom_histogram(bins = length(unique(bal_posts_clean$created_yearMonth) ),
                   fill = "blue", color = "black") + 
    ylab(label = "posts_count") ) + scale_fill_brewer(palette = "Dark2")

sub <- filter(bal_posts_clean, created_date > "2015-07-31")
sub <- sub[c("created_yearMonth", "likes_count",
                         "comments_count", "shares_count")]
gb_sub <- group_by(sub, created_yearMonth)
agg <- summarize(gb_sub, 
                 likes_mean=mean(likes_count),
                 comments_mean=mean(comments_count),
                 shares_mean=mean(shares_count),
                 mean_posts=n()/length(unique(created_yearMonth)),
                 n_YearMonths=length(unique(created_yearMonth)),
                 n_posts=n()
                 )
agg
```
There is high variability in number of posts per month. On average 24 posts are made per month exluding the first 4 months.


##

```{r, include=F}
sub <- bal_posts_clean[c("created_yearMonth", "type", "likes_count",
                         "comments_count", "shares_count")]

gb_sub <- group_by(sub, created_yearMonth, type)
agg <- summarize(gb_sub, 
                 likes_mean=mean(likes_count),
                 comments_mean=mean(comments_count),
                 shares_mean=mean(shares_count),
                 likes_total=sum(likes_count),
                 comments_total=sum(comments_count),
                 shares_total=sum(shares_count),
                 n_counts=n()
                 )

(p1 <-
    ggplot(data = agg,
           aes(x = created_yearMonth, y = value,
               color = variable, group = variable) ) +
    geom_line() +
    ylab(label = "posts_count") ) + scale_fill_brewer(palette = "Dark2")
```

## Post type by weekday

```{r, echo=F}
(p2 <- 
  ggplot(data = sub, 
         aes(x = created_weekday, fill = type) ) + 
  geom_histogram(stat="count") + 
  ylab(label = "posts_count") + scale_fill_brewer(palette = "Dark2")
)  
```

Link posts are slightly more common toward the end of the week. There are more link posts on Sunday than Saturday. 34.1% of link posts contain "vicpol" in the link.


## Comments and likes by post age

```{r, echo=F}
p3 <- 
  ggplot(data = bal_posts_clean, 
         aes(x = likes_count, y = comments_count, 
             shares_count = shares_count, type = type, id = id, 
             color = -post_age)) + 
  geom_point(alpha=.5) + scale_fill_brewer(palette = "Dark2")
plotly::ggplotly(p3)
```

Newer posts appear to receive more comments than older posts. Older posts may or may not receive for likes (hard to tell at this color and transparency)


## Shares and likes by type

```{r, echo=F}
p4 <- 
  ggplot(data = bal_posts_clean, 
         aes(x = likes_count, y = shares_count, 
             shares_count = shares_count, id = id, 
             color = type)) +
  geom_point(alpha=.5) + scale_fill_brewer(palette = "Dark2")
plotly::ggplotly(p4)
```

Extreme values tend to be photos. Extreme values tend to be liked or shared, but not both. Note that photos are the most common posts followed by links.


## Like by year-month across type of post

```{r, echo=F}
bal_posts_clean$type <- 
  factor(bal_posts_clean$type, levels = 
           rev(c("photo", "link", "video", "status", "music") ) )
(p5 <-
    ggplot(data = bal_posts_clean, 
       aes(x = created_yearMonth, y = likes_count, fill = type)) +
    geom_bar(stat = 'identity') + scale_fill_brewer(palette = "Dark2") 
)
```

Photos followed by links generate the most likes. Note that the y axis is likes and not likes/post.


### Missingness and pair-wise scatterplot

```{r, echo=F}
(p0 <- visdat::vis_dat(bal_posts_clean))
```

The data is mostly present. `story, message, link` are the variables by descending missingness.

```{r, echo=F}
sub <- 
  bal_posts_clean[c("created_date", "created_weekday", 
                    "likes_count", "comments_count", "shares_count", "type")]
(p00 <- GGally::ggpairs(data = sub))
```



### Metadata

page: eyewatchballarat
pull time: 1.04 min
raw dim: 952 obs x 11 var 
clean dim: 95 2obs x 26 var 
pull time: 5:30pm 16 May 2018
```{r, eval=F, include=F}
library(Rfacebook)

token <- "EAACEdEose0cBAPTWv9SPehGkMUhTbWCcBYkTi2FNvhAlETSACg0ZCrTxK94v7QSLXWJQxniD8aVkHob0541svS4WOVpKv8O1PABqXXtQWGusAMPdLZBYYRjo6MlOGVUHQQQCOEaoPZAoRWDXbKT8d7lBSId2mBvxOmqxUHJIbKE8UIObWPDOYDaXta9oxAZD" 
# temp token, will need to gen each time, see:
# https://developers.facebook.com/tools/explorer/?method=GET&path=me%3Ffields%3Did%2Cname&version=v3.0)

bal <- "eyewatchballarat"

start_time <- Sys.time()
bal_posts_raw <- as_tibble(getPage(bal, token, n=1000, feed = TRUE)) 
end_time <- Sys.time()

end_time - start_time #1.04 min, 952 x 11. ballarat, 5:30pm 16 May 2018
save(bal_posts_raw, file = "data/bal_posts_raw.rda")
```
