---
title: "Eyewatch facebook posts -- CSN station comparison"
author: "Nicholas Spyrison"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    self_contained: true
---
```{r options, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 8,
  fig.width = 12,
  fig.align = "center",
  cache = FALSE)

library("readr")
library("dplyr")
library("tibble")
library("tsibble")
library("ggplot2")
library("lubridate")
library("plotly")

load("../3_staged_data/bound_data_2019-04-09.rda")
dat <- bound_data
```

# Intro 

A cleaned, filtered data set of `r round(nrow(dat))` facebook posts accross `r round(ncol(dat))` variables for all 55 police stations pages was exported from Hootsuite. The data ranges from `r round(min(dat$Date))` through `r round(max(dat$Date))`. Posts before 18 November 2017 are not available on the Hootsuite platform. The below graphs were made from made from this data.

```{r, eval=T,include=F}
sub <- dat[c("Station", "YearMonth", "Reactions", "Comments",
                  "Shares", "Engagements")]
agg <- summarize(.data = sub,
                 CountStaions     = length(unique(Station)),
                 CountYearMonths  = length(unique(YearMonth)),
                 MeanPostsPm      = n()/length(unique(YearMonth)),
                 MeanReactionsPp  = mean(Reactions),
                 MeanSharesPp     = mean(Shares),
                 MeanCommentsPp   = mean(Comments),
                 MeanEngagementPp = mean(Engagements)
)
```



## Engagement by time

Something looks off with the Post Type after 2018/07. I will dig into this.

```{r, echo=F}

ggplot(data = dat, aes(x = YearMonth)) +
   geom_bar(aes(y = Engagements/1000, fill = PostType), stat = 'identity') + 
   ggtitle("Total Engagements vs time (by post type)") +
   xlab("Time (each bar is 1 month)") + ylab("'000x Engagments") +
   scale_fill_brewer(palette = "Dark2") +
   facet_grid(IsCsnStation~.)
```

This don't control for the post volume, let's look at Engagements per post.

```{r, echo=F}
agg1 <-
  dat %>%
  group_by(YearQuarter, PostType, IsCsnStation) %>%
  summarise (Engagements = sum(Engagements),
             Posts = n(),
             Engagements_pp = sum(Engagements)/n())

ggplot(data = agg1) +
  geom_bar(aes(x=YearQuarter, y=Engagements_pp, fill = PostType), stat = 'identity') +
  facet_grid(IsCsnStation~.)+
  ggtitle("Engagements per Post vs time (by post type)") +
   xlab("Time (each bar is 1 Quarter)") + ylab(" Engagments per Post") +
   scale_fill_brewer(palette = "Dark2")
```


CSN vs non-CSN are clearly very different, let's focus on CSN data below.

```{r CSN}
filt_dat <- filter(dat, IsCsnStation=="CSN")
ggplot(data = filt_dat, aes(x = YearMonth)) +
   geom_bar(aes(y = Engagements/1000, fill = PostType), stat = 'identity') + 
   ggtitle("Total Engagements vs time (by post type)") +
   xlab("Time (each bar is 1 month)") + ylab("'000x Engagments") +
   scale_fill_brewer(palette = "Dark2") +
   facet_grid(IsCsnStation~.)


#### LOOKING INTO Post type as, it still looks off; is it in the raw data?
#thinking: did videos turn into events, offseting the level somehow?

```