---
title: "CSN questions 1 & 2"
author: "Nicholas Spyrison"
date: '15 August 2019' # '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    self_contained: true
---

```{r options, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 8,
  fig.width = 12,
  fig.align = "center",
  cache = FALSE)

library("tidyverse")
library("plotly")

load("../3_staged_data/bound_data_2019-04-09.rda")
# str(bound_data) # ~39k x 39 var
dat <- bound_data %>% 
  filter(IsCsnStation == "CSN") # ~10k x 39 var 
dat$PostMessageLen <- nchar(dat$PostMessage)
dat$Hour <- as.integer(dat$Hour)
dat$Station <- as.factor(as.character(dat$Station)) # relevel
dat <- rename(dat, 
              `Post Type` = PostType,
              `Year Month` = YearMonth)
### Create 10 end quantiles
n = .1
dat$`Is Top` <- ifelse(dat$Engagements > quantile(dat$Engagements, prob = 1-n), T, F)
dat$`Is Bottom` <- ifelse(dat$Engagements < quantile(dat$Engagements, prob = n), T, F)

# ### EDA
# high <- subset(dat, Engagements > quantile(Engagements, prob = 1-n))
# low  <- subset(dat, Engagements < quantile(Engagements, prob = n)) 
# subset_vars1 <- c(2,3,5,7,30,38,39,40)
# dat_sub1 <- dat[, subset_vars1]
# high_sub <- high[, subset_vars1]
# GGally::ggpairs(high_sub) 
# summary(high_sub)
# summary(dat_sub1)
```

# Overview 

Answering the following questions with a paragraph and relevant graphs as prompted from Kathryn's 13 August 2019 email:
1. What types of posts had the highest and lowest engagements?
2. Was there a significant increase in engagement over the evaluation period?
    + If so, did the control, non-CSA sites increase too?
    
The scope of the data is facebook posts for the CSN stations between December 2017 and March 2019.

## High and low engagement

We are trying to identify or characterize where high/low (top and bottom 10%) engagement of posts come from rather than predict the engagements, because of this we tried to identify the demographic dimensions that show contrast, namely: date, day of the week, station, and post type. 

High engagement posts tend to be slightly newer,(1) though they are less sensitive to weekly fluctuations than the number of posts.(2) High engagement posts are disproportionately high for video and link post types.(3) Station is also a strong indicator of high engagement with Latrobe and Cardinia receiving a larger number, and Greater Dandenong and Greater Shepparton receiving a smaller share proportionately.(4)
```{r, eval=T, include=F}
##### High engagement
### Date density
dat %>%
 group_by(`Is Top`) %>%
 summarise(MeanDate = mean(Date)) -> agg1

gg1 <- ggplot(dat) +
  geom_density(alpha = .4, mapping = aes(x = Date, fill = `Is Top`)) +
  geom_vline(agg1, mapping = 
               aes(xintercept = MeanDate, linetype = `Is Top`)) +
  scale_fill_brewer(palette = "Dark2")

### Posts by Weekday
dat %>%
  group_by(`Is Top`, Weekday) %>%
  summarise(Posts = n()) -> agg2

gg2 <- ggplot(agg2, mapping = aes(x = Weekday, y = Posts, fill = `Is Top`)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Dark2")
# not too extreme

### Posts by Station
dat %>%
  group_by(`Is Top`, Station) %>%
  summarise(Posts = n()) -> agg3
agg3$Station <- factor(agg3$Station, 
                       levels = agg3$Station[c(6,5,3,7,12,10,1,9,8,2,4,11)])

gg3 <- ggplot(agg3, mapping = aes(x = Station, y = Posts, fill = `Is Top`)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1)) +
  scale_fill_brewer(palette = "Dark2")

### Posts by Type
dat %>%
  group_by(`Is Top`, `Post Type`) %>%
  summarise(Posts = n()) -> agg4
agg4$`Post Type` <- factor(agg4$`Post Type`, levels = agg4$`Post Type`[c(2,3,4,1,5)])

gg4 <- ggplot(agg4, mapping = aes(x = `Post Type`, y = Posts, fill = `Is Top`)) +
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Dark2")

```

**High engagement posts vs remainder**
```{r, echo=F}
cowplot::plot_grid(gg1, gg2, gg3, gg4, 
                     ncol = 2, nrow = 2,
                     labels = c(1,2,3,4))
```

Low engagement posts, surprisingly, also tend to be slightly newer,(5) they tend to be proportional to the number of posts on the weekdays, but less likely on Saturday.(6) Low engagement posts have a high proportion of status and photo post types, while link seems to have few such posts.(7) Station is a decent indicator of low engagement with Whittlesea has the largest proportion, while, Geelong and Wyndham rarely have low engagement.(8)

```{r, eval=T, include=F}
##### Low engagement
### Date density
dat %>%
 group_by(`Is Bottom`) %>%
 summarise(MeanDate = mean(Date)) -> agg5

gg5 <- ggplot(dat) +
  geom_density(alpha = .4, mapping = aes(x = Date, fill = `Is Bottom`)) +
  geom_vline(agg5, mapping = 
               aes(xintercept = MeanDate, linetype = `Is Bottom`)) +
  scale_fill_brewer(palette = "Dark2")

### Posts by Weekday
dat %>%
  group_by(`Is Bottom`, Weekday) %>%
  summarise(Posts = n()) -> agg6

gg6 <- ggplot(agg6, mapping = aes(x = Weekday, y = Posts, fill = `Is Bottom`)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Dark2")
# not too extreme

### Posts by Station
dat %>%
  group_by(`Is Bottom`, Station) %>%
  summarise(Posts = n()) -> agg7
agg7$Station <- factor(agg7$Station, 
                       levels = agg7$Station[c(6,5,3,7,12,10,1,9,8,2,4,11)])

gg7 <- ggplot(agg7, mapping = aes(x = Station, y = Posts, fill = `Is Bottom`)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 40, hjust = 1)) +
  scale_fill_brewer(palette = "Dark2")

### Posts by Type
dat %>%
  group_by(`Is Bottom`, `Post Type`) %>%
  summarise(Posts = n()) -> agg8
agg8$`Post Type` <- factor(agg8$`Post Type`, levels = agg8$`Post Type`[c(2,3,4,1,5)])

gg8 <- ggplot(agg8, mapping = aes(x = `Post Type`, y = Posts, fill = `Is Bottom`)) +
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Dark2")
```

**Low engagement posts vs remainder**
```{r, echo=F}
cowplot::plot_grid(gg5, gg6, gg7, gg8, 
                     ncol = 2, nrow = 2,
                     labels = c(5,6,7,8))
```

## Engagement over time

There is a strong seasonal effect on the mean number of engagements with a relatively weak increase month after month at a rate of about 3.75 engagements per post year over year.

```{r, eval=T, include=F}
dat %>%
  group_by(`Year Month`) %>%
  summarise(Posts = n(),
            `Mean Engagements` = mean(Engagements)) -> agg9

gg9 <- ggplot(agg9, mapping = aes(x = `Year Month`, y = `Mean Engagements`)) +
  geom_bar(stat = "identity") + 
  geom_smooth(se=F) +
  geom_smooth(method = "lm", se=F, color="black", formula = y ~ x, ) +
  scale_fill_brewer(palette = "Dark2")

agg9$YearMonthNum <- 1:nrow(agg9)
lm(`Mean Engagements` ~ YearMonthNum, data = agg9)
.312 * 12 # Year over year
```

**Engagement over time -- CSN stations**
```{r, echo=F}
cowplot::plot_grid(gg9, labels = 9)
```

Comparing the CSN stations with non-CSN stations we notice far less of seasonal effect which is shocking, we also notice a steep drop in mean engagements over time. The decrease is less pronounced after the initial drop.(10) We plot March 2018 to March 2019 for clearer comparison, finding that non-CSN stations decrease at a rate of about -1.0 engagements per post year over year.
```{r, eval=T, include=F}
notCSN <- bound_data %>% 
  filter(IsCsnStation == "not CSN") # ~29k x 39 var 

notCSN <- rename(.data = notCSN, `Year Month` = YearMonth)

notCSN %>%
  group_by(`Year Month`) %>%
  summarise(Posts = n(),
            `Mean Engagements` = mean(Engagements)) -> agg10

gg10 <- ggplot(agg10, mapping = aes(x = `Year Month`, y = `Mean Engagements`)) +
  geom_bar(stat = "identity") + 
  geom_smooth(se=F) +
  geom_smooth(method = "lm", se=F, color="black", formula = y ~ x, ) +
  scale_fill_brewer(palette = "Dark2")

agg10$YearMonthNum <- 1:nrow(agg10)
lm(`Mean Engagements` ~ YearMonthNum, data = agg10)
-1 # Year over year from march to march

### march to march
agg11 <- agg10[-1:-3, ]

gg11 <- ggplot(agg11, mapping = aes(x = `Year Month`, y = `Mean Engagements`)) +
  geom_bar(stat = "identity") + 
  geom_smooth(se=F) +
  geom_smooth(method = "lm", se=F, color="black", formula = y ~ x, ) +
  scale_fill_brewer(palette = "Dark2")

gg11

1 * 12 # Year over year
```

**Engagement over time -- non-CSN stations**

```{r, echo=F}
cowplot::plot_grid(gg10, gg11, ncol = 2, labels = c(10, 11))
```