---
title: "Eyewatch facebook posts -- via Hootsuite"
author: "Nicholas Spyrison"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    self_contained: true
---
```{r options, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 8,
  fig.width = 12,
  fig.align = "center",
  cache = FALSE)

library("readr")
library("dplyr")
library("tibble")
library("tsibble")
library("ggplot2")
library("lubridate")
library("plotly")

load("../3_staged_data/bound_data_2019-04-09.rda")
dat <- bound_data
```

A cleaned, filtered data set of `r round(nrow(dat))` facebook posts accross `r round(ncol(dat))` variables for all 55 police stations pages was exported from Hootsuite. The data ranges from `r round(min(dat$Date))` through `r round(max(dat$Date))`. Posts before 18 November 2017 are not available on the Hootsuite platform. The below graphs were made from made from this data.

```{r, eval=T,include=F}
sub <- dat[c("Station", "YearMonth", "Reactions", "Comments",
                  "Shares", "Engagements")]
agg <- summarize(.data = sub,
                 CountStaions     = length(unique(Station)),
                 CountYearMonths  = length(unique(YearMonth)),
                 MeanPostsPm      = n()/length(unique(YearMonth)),
                 MeanReactionsPp  = mean(Reactions),
                 MeanSharesPp     = mean(Shares),
                 MeanCommentsPp   = mean(Comments),
                 MeanEngagementPp = mean(Engagements)
)
```

## Posts by time

```{r, echo=F}
(g1 <-
  ggplot(data = dat, aes(x = `Date`, fill = PostType) ) +
  geom_histogram(bins = length(unique(dat$YearMonth))) +
  ggtitle("Post vs Time (by post type)") +
  ylab("Number of Posts") + xlab("Time (each bar is 1 month)") +
  scale_fill_brewer(palette = "Dark2") 
)
```

## Engagement by time


```{r, echo=F}

(g2 <- ggplot(data = dat, aes(x = YearMonth)) +
   geom_bar(aes(y = Engagements/1000, fill = PostType), stat = 'identity') + 
   ggtitle("Total Engagements vs time (by post type)") +
   xlab("Time (each bar is 1 month)") + ylab("1000x Engagments") +
   scale_fill_brewer(palette = "Dark2"))
```

This don't control for the post volume, let's look at Engagements per post.

```{r, echo=F}
agg1 <-
  dat %>%
  group_by(YearQuarter, PostType) %>%
  summarise (Engagements = sum(Engagements),
             Posts = n(),
             Engagements_pp = sum(Engagements)/n())

ggplot(data = agg1) +
  geom_bar(aes(x=YearQuarter, y=Engagements_pp, fill = PostType), stat = 'identity') +
  ggtitle("Engagements per Post vs time (by post type)") +
   xlab("Time (each bar is 1 month)") + ylab(" Engagments per Post") +
   scale_fill_brewer(palette = "Dark2")
```



<!-- ## Mean engagements trends by time across post type -->

<!-- Do we see engagement across post type changing over time? -->

<!-- ```{r, include=F} -->
<!-- # (g2 <- ggplot(dat,  -->
<!-- #               aes(x = `Date`, y = Engagements, fill = PostType ) ) + -->
<!-- #    geom_smooth() + scale_fill_brewer(palette = "Dark2") + xlab("Time") +  -->
<!-- #    coord_cartesian(ylim = c(0, 100))  -->
<!-- # ) -->
<!-- ``` -->

## Engagement by post type across post age

Of the bottom 95% of posts by engagement.

```{r, echo=F}
p3 <- ggplot(data = dat, 
             aes(x = PostType, y = Engagements, 
                 color = -DaysOld, id = PagePostID)) + 
  geom_point(alpha = .4, position = "jitter") + 
  scale_y_continuous(limits = c(0, quantile(dat$Engagements, .95)) ) 
#plotly::ggplotly(p3)
p3
```

## High engagment posts: engagement by post type across post age 

Taking a closer look at high engagement posts. Looking at the same graph for only the top 2% of engaged posts.

```{r, echo=F}
p4 <- ggplot(data = dat, 
             aes(x = PostType, y = Engagements, 
                 color = -DaysOld, id = PagePostID)) + 
  geom_point(alpha = .6, position = "jitter") + 
  scale_y_continuous(limits = c(
    quantile(dat$Engagements, .98),
                         max(dat$Engagements)*1.01 ) 
  ) + scale_fill_brewer(palette = "Dark2")
#plotly::ggplotly(p4)
p4
```

### Appendix

All times are listed in GMT. `story` is NA if there is no message with a post, other than that the data has low missingness.

```{r, echo=F}
sub <- dat[c("YearMonth", "Reactions_Anger", "Reactions_Haha",
                   "Reactions_Like","Reactions_Love", "Reactions_Sorry",
                   "Reactions_Wow", "Reactions", "Comments","Shares", 
                   "Engagements")]
sub_samp <- sample_n(sub, 1000)
(a2 <- GGally::ggpairs(data = sub_samp))
```


```{r, echo=F}
a3sub <- 
  dat[c("Station", "Date","Weekday", "DaysOld", "PostType",
             "Reactions_Anger", "Reactions_Haha", "Reactions_Like",
             "Reactions_Love", "Reactions_Sorry", "Reactions_Wow", "Reactions",
             "Comments","Shares", "Engagements")]
summary(a3sub)
```